---
title: "Automatic and Manual Return of the Timings"
author: Jonathan Berrisch
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: no
    toc: no
vignette: >
  %\VignetteIndexEntry{Autoreturn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Assign a Custom Name to the Timings Object

If you want to give the dataframe a different name, you can pass that name to the constructor:

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer("mytimes");
  timer.tic();
  std::string s;
  s.reserve(1048576);
  timer.toc();
  return(0);
}',
  depends = "rcpptimer"
)

mem()

print(mytimes)
```

```{r, eval = TRUE, echo = FALSE}
rm(list = ls())
```

## Manually Handling the Results

It may be necessary for you to handle the resulting DataFrame yourself instead of letting `rcpptimer` write it to the global environment. 

This issue actually consists of two elements:

- Turn off the autoreturn feature
- Handle the results yourself

First, to turn off the autoreturn feature, set the `autoreturn` variable of the `timer` class to `false`. This will prevent the `timer` from writing the results to the global environment. 

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer;
  timer.autoreturn = false;
  timer.tic("mem");
  std::string s;
  s.reserve(1048576);
  timer.toc("mem");
  return(0);
}',
  depends = "rcpptimer"
)

mem()

print(ls())
```

Now, the results are not written to the global environment. Instead, you can access them through the `timer` object using the `stop()` method.

```{r, eval = TRUE}
Rcpp::cppFunction('
DataFrame mem()
{
  Rcpp::Timer timer;
  timer.autoreturn = false;
  timer.tic("mem");
  std::string s;
  s.reserve(1048576);
  timer.toc("mem");
  DataFrame times = timer.stop();
  return(times);
}',
  depends = "rcpptimer"
)

mem()
```

Its also possible to use `stop()` *and* let rcpptimer pass the results automatically. In the above example, we just set autoreturn to true (the default). 

In the above example, `mem()` returns the results *and* rcpptimer writes them as `times` to the global environment.

## Considerations for package authors

```{r, eval = TRUE, echo = FALSE}
rm(list = ls())
```

As a package author, you may want to parameterize the autoreturn feature so that your users can decide whether they want the results to be written to the global environment and what name should be assigned. Below is an example of how to do this.

```{r, eval = TRUE}
Rcpp::cppFunction('
void mem(bool return_timings = false,
         const char* timings_name = "times")
{
  Rcpp::Timer timer(timings_name);
  timer.autoreturn = return_timings;
  timer.tic("mem");
  std::string s;
  s.reserve(1048576);
  timer.toc("mem");
}',
  depends = "rcpptimer"
)

mem(return_timings = FALSE)
ls()
mem(return_timings = TRUE)
ls()
mem(return_timings = TRUE, timings_name = "my_times")
ls()
```
